% Figures


```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(cowplot)
library(patchwork)
library(ggrepel)
library(viridis)
library(lme4) 
library(emmeans)
library(LaplacesDemon)
library(car)
```

## Figure 1

### Ethiopia Map

```{r}
library(rnaturalearth)
library(rnaturalearthhires)
library(sf)


ETH <- ne_countries(
  country = "Ethiopia",
  returnclass = "sf"
)

region <- sf::st_read("data/ethiopiaregion/Eth_Region_2013.shp")
zone <- sf::st_read("data/ethiopia-zone/Eth_Zone_2013.shp")
district <- sf::st_read("data/ethiopiaworeda/Eth_Woreda_2013.shp")
```


```{r}

ETH_REGIONS <- region %>%
  filter(REGIONNAME %in% c("SNNPR", "Oromia"))

survey$zone <- plyr::revalue(
  survey$zone,
  c(
    "Ilu AbaBora" = "Ilubabor",

    "West Welega" = "West Wellega"
  )
)

zone_names <- survey %>%
  select(zone) %>%
  unique()

ETH_ZONE <- zone %>%
  filter(ZONENAME %in% c(
    "Jimma",
    "West Wellega",
    "Sidama",
    "Sheka",
    "Keffa",
    "Bench Maji",
    "Bale",
    "Gedio",
    "Ilubabor"
  ))

district_names <- survey %>%
  select(district) %>%
  unique()

survey$district <- plyr::revalue(survey$district, c(
  "Aira" = "Ayira",
  "Aletawondo" = "Aleta Wendo",
  "Anderecha" = "Anderacha",
  "Debub B" = "Debub Bench",
  "Delo-Menna" = "Mena",
  "Di/Zuria" = "Dila Zuria",
  "Gurafarda" = "Gurafereda",
  "Harena" = "Harena Buluk",
  "Mettu" = "Metu Zuria",
  "Shebe-Sombo" = "Shebe Sambo",
  "Shebedino" = "Shebe Dino",
  "Sheko" = "Sheka",
  "Wonago" = "Wenago",
  "Y/chefe" = "Yirgachefe",
  "Yayo" = "Yayu",
  "Gomma" = "Goma"
))
district_names <- district_names$district


ETH_DISTRICTS <- district %>%
  filter(WOREDANAME %in% district_names)
```




```{r fig.height=3, fig.width=7, message=FALSE, warning=FALSE}
survey <- read_csv("data/survey_clean.csv")

```


```{r eval=FALSE, include=FALSE}
library(ggsflabel)
library(plotly)
library(ggsn)
ETH2 <- st_transform(ETH, 31983)

map_district <- ggplot(ETH) +
  # geom_sf(data = ETH, fill = "grey92", size = 0.3, color = "white") +

  geom_sf(data = ETH_ZONE, color = NA, size = 0) +
  geom_sf(data = ETH_REGIONS, alpha = 0.1, fill = "grey90", color = "grey30", aes(linetype = REGIONNAME), size = 0.9) +
  geom_sf(data = ETH_DISTRICTS,  color = "black", size = 0.4) +
  coord_sf(datum = NA) +
  theme_map() +
  theme(legend.position = "right") +
  scale_fill_brewer(type = "qual", palette = "Set3") +
  geom_sf_text_repel(
    dat = ETH_DISTRICTS, aes(label = WOREDANAME), size = 4
  ) +
  scale_color_viridis() +
  geom_point(data = survey, aes(lon, lat, color = altitude), size = 1, shape = 16)
  # guides(colour = guide_legend(override.aes = list(size = 4))) +
  # labs(
  # x = "Longitude", y = "Latitude", fill = "Zone", linetype = "Region",
  # ) +


  ggsave("figs/map_districts.png", width = 15, dpi = 300)
```


```{r eval=FALSE, include=FALSE}
plot_grid(map_district, labels = LETTERS, ncol = 1, align = "v")
ggsave("figs/Figure1.png", width = 10)
```



## Figure 2

### Inc-Sev relationship



## Figure 3

### Incidence vs zone


```{r}
p_inc_alt_zone <- survey %>%
  ggplot(aes(altitude, logit(inc/100), color = zone, group = zone)) +
  geom_point() +
  geom_smooth(method = lm, se = F) +
  theme_minimal_grid(font_size = 10) +
  theme(legend.position = "none") +
  scale_color_brewer(type = "qual", palette = "Set3")+
    labs(x = "Altitude (m)", y = "Logit Incidence (proportion)")

```


### Severity vs zone


```{r}
p_sev_alt_zone <- survey %>%
  ggplot(aes(altitude, logit(sev2/100), color = zone, group = zone)) +
  geom_point() +
  geom_smooth(method = lm, se = F) +
  theme_minimal_grid(font_size = 10) +
  theme(legend.position = "none") +
  scale_color_brewer(type = "qual", palette = "Set3")+
    labs(x = "Altitude (m)", y = "Logit Severity (Proportion)")

```


##### Plot slopes  inc by zone


#### Effect on incidence


```{r}
m_alt_inc_zone <- lmer(logit(inc/100 ) ~ altitude * zone + (1 | district), survey, REML = F)
summary(m_alt_inc_zone)
Anova(m_alt_inc_zone, type = "III")

zone_alt_inc <- data.frame(emtrends(m_alt_inc_zone, pairwise ~ zone, var = "altitude" ))



```





```{r}
p_slopes_inc_alt_zone <- zone_alt_inc %>% 
  ggplot(aes(reorder(emtrends.zone, emtrends.altitude.trend),emtrends.altitude.trend, color = emtrends.zone))+
  geom_point(size =3)+
  coord_flip()+
  theme_minimal_vgrid(font_size = 11)+
  theme(legend.position = "none")+
   scale_color_brewer(type = "qual", palette = "Set3")+
  labs(x = "", y = "Slope estimates")+
  geom_errorbar(aes(ymin = emtrends.lower.CL, ymax = emtrends.upper.CL),
                width =0.2, size =1)
```


##### Plots slopes sev by zone


```{r}
m_alt_sev_zone <- lmer(logit(sev2 ) ~ altitude * zone + (1 | district), data = survey, REML = F)
Anova(m_alt_sev_zone)
zone_alt_sev <- data.frame(emtrends(m_alt_sev_zone, pairwise ~ zone, var = "altitude" ))
```



```{r}
p_slopes_sev_alt_zone <- zone_alt_sev %>% 
  ggplot(aes(reorder(emtrends.zone, emtrends.altitude.trend),emtrends.altitude.trend, color = emtrends.zone))+
  geom_point(size =3)+
  coord_flip()+
  theme_minimal_vgrid(font_size = 11)+
  theme(legend.position = "none")+
   scale_color_brewer(type = "qual", palette = "Set3")+
  labs(x = "", y = "Slope estimates")+
  geom_errorbar(aes(ymin = emtrends.lower.CL, ymax = emtrends.upper.CL),
                width =0.2, size =1)

```

##### Panel of plots 

```{r}
((p_inc_alt_zone  / p_sev_alt_zone ) | (p_slopes_inc_alt_zone / p_slopes_sev_alt_zone)) +
  plot_annotation(tag_levels = "A") 
  ggsave("figs/Figure3.png", width = 8, height = 7)
```


## Figure 4

### Cropping system by zone

```{r}
survey$cropping_system <- factor(survey$cropping_system, levels = c("Forest", "Semiforest", "Garden", "Plantation"))

csystem <- survey %>% 
  tabyl(zone, cropping_system) %>% 
  gather(cropping_system, name, 2:5)

csystem$cropping_system <- factor(csystem$cropping_system, levels = c("Forest", "Semiforest", "Garden", "Plantation"))

p_cropping <- csystem %>% 
  ggplot(aes(zone, name, fill = cropping_system))+
  geom_col(position = "fill", color = "black", size =0.3)+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(fill = "",
       x = "",
       y = "Proportion")+
  scale_fill_brewer(palette = "Greens", direction = -1)
p_cropping
```

### Cropping system by altitude

```{r}

survey$cropping_system <- factor(survey$cropping_system, levels = c("Forest", "Semiforest", "Garden", "Plantation"))
p_cropping_altitude <- survey %>% 
  dplyr::select(altitude, cropping_system, inc, sev2) %>% 
  ggplot(aes(altitude, sev2, color = cropping_system, group = cropping_system))+
  geom_point(size = 3, shape = 16)+
 # geom_smooth(method = "lm", se = F )+
  theme_minimal_grid(font_size=11)+
  theme(legend.position = "top")+
  labs(color = "",
       x = "Altitude (m)",
       y = "Severity (%)")+
  scale_color_brewer(palette = "Greens", direction = -1)
p_cropping_altitude
  
```


### Farm management by zone

```{r}

farm_man <- survey %>% 
  tabyl(zone, farm_management) %>% 
  gather(farm_management, name, 2:5) 


farm_man$farm_management <- factor(
farm_man$farm_management, levels = c("Intensive", "Moderate", "Minimal", "Unmanaged"))

p_farm <- farm_man %>% 
  ggplot(aes(zone, name, fill = farm_management))+
  geom_col(position = "fill", color = "black", size =0.3)+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(fill = "",
       x = "",
       y = "Proportion")+
  scale_fill_brewer(palette = "Blues", direction = -1)

  
```



### Farm management by altitude


```{r}



survey$farm_management <- factor(
survey$farm_management, levels = c("Intensive", "Moderate", "Minimal", "Unmanaged"))
p_farm_man_altitude <- survey %>% 
  dplyr::select(altitude, farm_management, inc, sev2) %>% 
  ggplot(aes(altitude, sev2, color = farm_management, group = farm_management))+
  geom_point(size = 3, shape = 16)+
 # geom_smooth(method = "lm", se = F )+
  theme_minimal_grid(font_size=11)+
  theme(legend.position = "top")+
  labs(color = "",
       x = "Altitude (m)",
       y = "Severity (%)")+
scale_color_brewer(palette = "Blues", direction = -1)
p_farm_man_altitude
  
```

### Shade by zone


```{r}
library(janitor)
survey_shade <- survey %>% 
  tabyl(zone, shade) %>% 
  gather(shade, name, 2:4) 


survey_shade$shade <- factor(
survey_shade$shade, levels = c("Full shade", "Mid shade", "Sun"))

p_shade <- survey_shade %>% 
  ggplot(aes(zone, name, fill = shade))+
  geom_col(position = "fill", color = "black", size =0.3)+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(fill = "",
       x = "",
       y = "Proportion")+
  scale_fill_brewer(palette = "OrRd", direction = 1)
p_shade
```

### Shade by altitude


```{r}

survey$shade <- factor(
survey$shade, levels = c("Full shade", "Mid shade", "Sun"))

p_shade_altitude <- survey %>% 
  dplyr::select(altitude, shade, inc, sev2) %>% 
  ggplot(aes(altitude, sev2, color = shade, group = shade))+
  geom_point(size = 3, shape = 16)+
 # geom_smooth(method = "lm", se = F )+
  theme_minimal_grid(font_size=11)+
  theme(legend.position = "top")+
  labs(color = "",
       x = "Altitude (m)",
       y = "Severity (%)")+
  scale_color_brewer(palette = "OrRd", direction = 1)
p_shade_altitude
  
```


### Cultivar by zone

```{r}

survey_cult <-survey %>% 
  tabyl(zone, cultivar) %>% 
  gather(cultivar, name, 2:4) 
  

survey_cult$cultivar <- factor(

survey_cult$cultivar, levels = c("Improved", "Mixture", "Local", "Unmanaged"))

p_cult <- survey_cult %>% 
  ggplot(aes(zone, name, fill = cultivar))+
  geom_col(position = "fill", color = "black", size =0.3)+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(fill = "",
       x = "",
       y = "Proportion")+
  scale_fill_brewer(palette = "PuRd", direction = -1)
p_cult



```


### Cultivar by altitude


```{r}

survey$cultivar <- factor(

survey$cultivar, levels = c("Improved", "Mixture", "Local", "Unmanaged"))
p_cult_altitude <- survey %>% 
  dplyr::select(altitude, cultivar, inc, sev2) %>% 
  ggplot(aes(altitude, sev2, color = cultivar, group = cultivar))+
  geom_point(size = 3, shape = 16)+
 # geom_smooth(method = "lm", se = F )+
  theme_minimal_grid(font_size=11)+
  theme(legend.position = "top")+
  labs(color = "",
       x = "Altitude (m)",
       y = "Severity (%)")+
  scale_color_brewer(palette = "PuRd", direction = -1)
p_cult_altitude
  
```



### Compose figure

```{r}
library(patchwork)

 ((p_cropping_altitude / p_cult_altitude / p_farm_man_altitude / p_shade_altitude) |
  ((p_cropping / p_cult /p_farm / p_shade ) ))+ 
      plot_annotation(tag_levels = 'A') 
ggsave("figs/Figure4.png", width =8, height =10)

```

(p_cropping_altitude / p_cult_altitude / p_farm_man_altitude / p_shade_altitude))+ 




## Figure 5

### MCA plot

```{r}
library(FactoMineR)
library(factoextra)
survey_mca <- survey %>% 
 mutate(altitude2 = case_when(
    altitude < 1452 ~ "< 1452",
    altitude < 1685 ~ "1452 to 1685",
    TRUE ~ "> 1685m"
  )) 
 
attach(survey_mca)
data_mca <- survey_mca %>% dplyr::select(altitude2, zone, cultivar, shade, cropping_system, farm_management)

head(data_mca)
cats <- apply(data_mca, 2, function(x) nlevels(as.factor(x))) # enumera as categorias
cats
res.mca <- MCA(data_mca, graph = FALSE)

p <- fviz_mca_var(res.mca,
  label = "var", repel = T,
  col.var = "contrib",
  # Avoid text overlapping (slow if many point)
  ggtheme = theme_minimal()
)
p + scale_color_viridis() +
  labs (title = "", fill = "Contribution")+
  ggsave("figs/Figure5.png", width =7, height =5)
```

